---
title: "Supplementary S2\nFunctional diversity is decoupled from phylogenetic diversity across plant communities, reflecting biogeographic and climatic history."
output: 
  word_document:
    reference_docx: ref.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("tidyverse")
library("mgcv")
library("performance")
library("dismo")
library("gbm")
library("googlesheets4")
library("tidyr")

```

```{r, warning=FALSE, message=FALSE, echo=FALSE}

gs4_auth()

authors <- read_sheet("https://docs.google.com/spreadsheets/XXXXXXXXXXXX", sheet = "Affiliations")

authors <- authors %>% 
  mutate(
  Affiliation1 = ifelse(!is.na(`Affiliation 1`), paste0(`Affiliation 1`, ", ", `Address 1`), NA),
  Affiliation2 = ifelse(!is.na(`Affiliation 2`), paste0(`Affiliation 2`, ", ", `Address 2`), NA),
  Affiliation3 = ifelse(!is.na(`Affiliation 3`), paste0(`Affiliation 3`, ", ", `Address 3`), NA)
  ) %>% 
  dplyr::select(Affiliation1, Affiliation2, Affiliation3, Name, Mail, ORCID)


# Get affiliations while keeping the original author order
affil <- gather(authors, key = "Level", value = "Affiliation", -c(Name, Mail, ORCID)) %>% 
  filter(stringr::str_detect(Level, "Affil")) %>% 
  arrange(match(Name, authors$Name))

affil_order <- data.frame(Affiliation = na.omit(unique(affil$Affiliation)),
                          Number = 1:length(na.omit(unique(affil$Affiliation))))

authors_numbers = affil %>% 
  right_join(affil_order %>% na.omit(), by = "Affiliation") %>% 
  group_by(Name) %>% 
  summarize(Affil = paste(Number, collapse = ",")) %>% 
  arrange(match(Name, authors$Name))


authorlist <- ""

for (i in 1:length(authors_numbers$Name)) {

  str <- paste0(authors_numbers[i, "Name"], "^", ifelse(i == 1, paste0(authors_numbers[i, "Affil"], ",\\*"), 
                                                        authors_numbers[i, "Affil"]), 
                       "^", ifelse(i == length(authors_numbers$Name), "",  ", "), sep = "")
  authorlist <- paste0(authorlist, str, sep = "")
}

affillist <- "\\* Corresponding author   \n   \n"

for (i in 1:length(affil_order$Affiliation)) {
  a <- paste0("^", affil_order[i, "Number"], "^", " ", affil_order[i, "Affiliation"], sep = "\n")
  affillist <- paste0(affillist, a)
}

```
`r authorlist`

<!-- ```{r, results='asis', echo=FALSE} -->
<!-- mycat <- function(text){ -->
<!--   cat(gsub(pattern = "\n", replacement = "  \n", x = text)) -->
<!-- } -->
<!-- mycat(affillist) -->
<!-- ``` -->

## Figure 2

### Figure 2 A

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEF ~ SES.RQEP + s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
mod1 <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-SES-RQEF-RQEP.Rds")

summary(mod1)

#check_model(mod1)

rm(list = ls())
```

## Figure 3

#### Model structure

```{r, echo=TRUE, eval=FALSE}
gbm.step(data = sPlot.data,
         gbm.x = "PC1",
                 "PC2",
                 "PC3",
                 "PC4",
                 "PC5",
                 "plot.size", "Plants recorded", #data bias
                 "sBiome", "is.forest", #Vegetation type
                 "stable.clim" #Longterm climate stability (LGM)
                 ),
         gbm.y = "SES.RQEP", # OR "SES.RQEF"
         family = "gaussian",
         tree.complexity = 5,
         learning.rate = 0.01,
         bag.fraction = 0.5,
         max.trees = 20000,
         silent = F)
```

#### Summary

```{r, echo = FALSE,comment=NA}

BRT.SES.FD <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/eve-calculations/BRT-calculation/04.BRT-results-SES.RQEF.RDS")

summary(BRT.SES.FD)

BRT.SES.PD <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/eve-calculations/BRT-calculation/04.BRT-results-SES.RQEP.RDS")

summary(BRT.SES.PD)

rm(list = ls())
```

## Figure 4

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEF ~ stable.clim + PC2 + PC3 + PC5 + plot.size + is.forest +
            s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}

mod.FD.complete <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/eve-calculations/GAM-expl/05.GAM_SES.RQEF-exp.RDS")

summary(mod.FD.complete)

#check_model(mod.FD.complete)

rm(list = ls())

```

### Figure 4 A

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEF ~ PC2 + PC3 + PC5 + plot.size + is.forest+ s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
mod.PD.SC <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF-exp_1.RDS")

summary(mod.PD.SC)

rm(list = ls())
```

### Figure 4 B

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEF ~ stable.clim + PC3 + PC5 + plot.size + is.forest +
            s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}

mod.FD.PC2 <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF-exp_2.RDS")

summary(mod.FD.PC2)

rm(list = ls())
```

### Figure 4 C

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEF ~ stable.clim + PC2 + PC5 + plot.size + is.forest +
            s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}

mod.FD.PC3 <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF-exp_3.RDS")

summary(mod.FD.PC3)

rm(list = ls())
```

### Figure 4 D

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEF ~ stable.clim + PC2 + PC3 + plot.size + is.forest +
            s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}

mod.FD.PC5 <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF-exp_4.RDS")

summary(mod.FD.PC5)

rm(list = ls())
```

### Figure 4 E

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEF ~ stable.clim + PC2 + PC3 + PC5 + is.forest +
            s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}

mod.FD.PS <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF-exp_5.RDS")

summary(mod.FD.PS)

rm(list = ls())
```

### Figure 4 F

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEF ~ stable.clim + PC2 + PC3 + PC5 + plot.size +
            s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}

mod.FD.for <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF-exp_6.RDS")

summary(mod.FD.for)

rm(list = ls())
```

### Figure 4 G

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEF ~ 1 + s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}

mod.FD.SP <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF-exp_7.RDS")

summary(mod.FD.SP)

rm(list = ls())
```

## Figure 5

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEP ~ PC1 + is.forest + s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}

mod.PD.complete <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/eve-calculations/GAM-expl/05.GAM_SES.RQEP-exp.RDS")

summary(mod.PD.complete)

#check_model(mod.PD.complete)

rm(list = ls())

```

### Figure 5 A

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEP ~ is.forest + s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
mod.PD.PC1 <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEP-exp_1.RDS")

summary(mod.PD.PC1)

rm(list = ls())
```

### Figure 5 B

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEP ~ PC1 + s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
mod.PD.for <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEP-exp_2.RDS")

summary(mod.PD.for)

rm(list = ls())
```

### Figure 5 C

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEP ~ 1 + s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
mod.PD.sp <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEP-exp_3.RDS")

summary(mod.PD.sp)

rm(list = ls())
```

## Figure S 1

### Figure S 1 A

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(FD ~ PD + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS1A <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-RQEF-RQEP.Rds")

summary(modS1A)

#check_model(modS1A)

rm(list = ls())
```

### Figure S 1 B

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(FD ~ PD.sqrt + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS1B <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-RQEF-RQEP.sqrt.Rds")

summary(modS1B)

#check_model(modS1B)

rm(list = ls())
```

### Figure S 1 C

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.FD ~ SES.PD.sqrt + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS1C <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-SES-RQEF-RQEP.sqrt.Rds")

summary(modS1C)

#check_model(modS1C)

rm(list = ls())
```

### Figure S 1 D

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(FD ~ 1 + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS1D <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-RQEF-RQEP.Smooth.Rds")

summary(modS1D)

#check_model(modS1C)

rm(list = ls())
```

## Figure S 7

### Figure S 7 A

#### Model structure

```{r, echo=TRUE, eval=FALSE}

indices <- c(
    "PDQ",
    "FDQ.ALL",
    "FDQ.SLA",
    "FDQ.HEIGHT",
    "FDQ.ROOT",
    "FDQ.SM",
    "FDQ.LDMC",
    "FDQ.N",
    "FDQ.P",
    "FDQ.CHRO.PA")

mgcv::gam(indices[i] ~ PDQ.PA + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
models <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/sPlotOpen-PA-GAMS.Rds")

blanks <- rep(c(' ', '\n'),3) # This will print five blank lines between plots.

AddBreak <- function() {
for (i in blanks)
  cat(i)
}

for(i in 1:length(models)) {
  print(summary(models[[i]]))
  AddBreak()
}

rm(list = ls())
```