---
title: "Supplementary S2\nFunctional diversity is decoupled from phylogenetic diversity across plant communities, reflecting biogeographic and climatic history."
output: 
  word_document:
    reference_docx: ref.docx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("tidyverse")
library("mgcv")
library("performance")
library("dismo")
library("gbm")
library("googlesheets4")
library("tidyr")

```

```{r, warning=FALSE, message=FALSE, echo=FALSE}

gs4_auth()

authors <- read_sheet("https://docs.google.com/spreadsheets/d/1ro2OU2A0lr9CzPqAcnNLic02aJIWZv2T_t_aJ3pU1sk/edit?usp=sharing", sheet = "Affiliations")

authors <- authors %>%
  mutate(
  Affiliation1 = ifelse(!is.na(`Affiliation 1`), paste0(`Affiliation 1`, ", ", `Address 1`), NA),
  Affiliation2 = ifelse(!is.na(`Affiliation 2`), paste0(`Affiliation 2`, ", ", `Address 2`), NA),
  Affiliation3 = ifelse(!is.na(`Affiliation 3`), paste0(`Affiliation 3`, ", ", `Address 3`), NA)
  ) %>%
  dplyr::select(Affiliation1, Affiliation2, Affiliation3, Name, Mail, ORCID)


# Get affiliations while keeping the original author order
affil <- gather(authors, key = "Level", value = "Affiliation", -c(Name, Mail, ORCID)) %>%
  filter(stringr::str_detect(Level, "Affil")) %>%
  arrange(match(Name, authors$Name))

affil_order <- data.frame(Affiliation = na.omit(unique(affil$Affiliation)),
                          Number = 1:length(na.omit(unique(affil$Affiliation))))

authors_numbers = affil %>%
  right_join(affil_order %>% na.omit(), by = "Affiliation") %>%
  group_by(Name) %>%
  summarize(Affil = paste(Number, collapse = ",")) %>%
  arrange(match(Name, authors$Name))


authorlist <- ""

for (i in 1:length(authors_numbers$Name)) {

  str <- paste0(authors_numbers[i, "Name"], "^", ifelse(i == 1, paste0(authors_numbers[i, "Affil"], ",\\*"),
                                                        authors_numbers[i, "Affil"]),
                       "^", ifelse(i == length(authors_numbers$Name), "",  ", "), sep = "")
  authorlist <- paste0(authorlist, str, sep = "")
}

affillist <- "\\* Corresponding author   \n   \n"

for (i in 1:length(affil_order$Affiliation)) {
  a <- paste0("^", affil_order[i, "Number"], "^", " ", affil_order[i, "Affiliation"], sep = "\n")
  affillist <- paste0(affillist, a)
}

```
`r authorlist`

```{r, results='asis', echo=FALSE}
mycat <- function(text){
  cat(gsub(pattern = "\n", replacement = "  \n", x = text))
}
mycat(affillist)
```

## Figure 2

### Figure 2 A

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEF ~ SES.RQEP + s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
mod1 <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-SES-RQEF-RQEP.bio.weigh.Rds")

summary(mod1)

#check_model(mod1)

rm(list = ls())
```

## Figure 3

#### Model structure

```{r, echo=TRUE, eval=FALSE}
gbm.step(data = sPlot.data,
         gbm.x = "PC1",
                 "PC2",
                 "PC3",
                 "PC4",
                 "PC5",
                 "Plants recorded", #data bias
                 "is.forest", #Vegetation type
                 "stable.clim" #Longterm climate stability (LGM)
                 ),
         gbm.y = "SES.RQEP", # OR "SES.RQEF"
         family = "gaussian",
         tree.complexity = 5,
         learning.rate = 0.01,
         bag.fraction = 0.5,
         max.trees = 20000,
         silent = F)
```

#### Summary

```{r, echo = FALSE,comment=NA}

BRT.SES <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.BRT-results-SES.RQEF.BW.RDS")

summary(BRT.SES[[1]])

summary(BRT.SES[[2]])

rm(list = ls())

BRT.rel.log <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.BRT-results-SES.BW.rel.log.RDS")

summary(BRT.rel.log)

rm(list = ls())
```

## Figure 4

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEF ~ stable.clim + PC2 + PC5 + 
            s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}

mod.FD.complete <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF.BWexp_5.RDS")

summary(mod.FD.complete)

#check_model(mod.FD.complete)

rm(list = ls())

```

### Figure 4 A

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEF ~ stable.clim + PC5 + s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
mod.PD.PC2 <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF.BWexp_7.RDS")

summary(mod.PD.PC2)

rm(list = ls())
```

### Figure 4 B

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEF ~ stable.clim + PC2 +  
            s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}

mod.FD.PC5 <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF.BWexp_8.RDS")

summary(mod.FD.PC5)

rm(list = ls())
```

### Figure 4 C

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEF ~ PC2 + PC5 +
            s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}

mod.FD.SC <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF.BWexp_6.RDS")

summary(mod.FD.SC)

rm(list = ls())
```

<!-- ### Figure 4 D -->

<!-- #### Model structure -->

<!-- ```{r, echo=TRUE, eval=FALSE} -->
<!-- mgcv::gam(SES.RQEF ~ stable.clim + PC2 + PC3 + plot.size + is.forest + -->
<!--             s(Longitude, Latitude, bs = "sos"), -->
<!--           family = "gaussian", -->
<!--           method = "REML", -->
<!--           data = data.wgs) -->
<!-- ``` -->

<!-- #### Summary -->

<!-- ```{r, echo = FALSE,comment=NA} -->

<!-- mod.FD.PC5 <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF-exp_4.RDS") -->

<!-- summary(mod.FD.PC5) -->

<!-- rm(list = ls()) -->
<!-- ``` -->

<!-- ### Figure 4 E -->

<!-- #### Model structure -->

<!-- ```{r, echo=TRUE, eval=FALSE} -->
<!-- mgcv::gam(SES.RQEF ~ stable.clim + PC2 + PC3 + PC5 + is.forest + -->
<!--             s(Longitude, Latitude, bs = "sos"), -->
<!--           family = "gaussian", -->
<!--           method = "REML", -->
<!--           data = data.wgs) -->
<!-- ``` -->

<!-- #### Summary -->

<!-- ```{r, echo = FALSE,comment=NA} -->

<!-- mod.FD.PS <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF-exp_5.RDS") -->

<!-- summary(mod.FD.PS) -->

<!-- rm(list = ls()) -->
<!-- ``` -->

<!-- ### Figure 4 F -->

<!-- #### Model structure -->

<!-- ```{r, echo=TRUE, eval=FALSE} -->
<!-- mgcv::gam(SES.RQEF ~ stable.clim + PC2 + PC3 + PC5 + plot.size + -->
<!--             s(Longitude, Latitude, bs = "sos"), -->
<!--           family = "gaussian", -->
<!--           method = "REML", -->
<!--           data = data.wgs) -->
<!-- ``` -->

<!-- #### Summary -->

<!-- ```{r, echo = FALSE,comment=NA} -->

<!-- mod.FD.for <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF-exp_6.RDS") -->

<!-- summary(mod.FD.for) -->

<!-- rm(list = ls()) -->
<!-- ``` -->

<!-- ### Figure 4 G -->

<!-- #### Model structure -->

<!-- ```{r, echo=TRUE, eval=FALSE} -->
<!-- mgcv::gam(SES.RQEF ~ 1 + s(Longitude, Latitude, bs = "sos"), -->
<!--           family = "gaussian", -->
<!--           method = "REML", -->
<!--           data = data.wgs) -->
<!-- ``` -->

<!-- #### Summary -->

<!-- ```{r, echo = FALSE,comment=NA} -->

<!-- mod.FD.SP <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF-exp_7.RDS") -->

<!-- summary(mod.FD.SP) -->

<!-- rm(list = ls()) -->
<!-- ``` -->

## Figure 5

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEP ~ PC1 + is.forest + s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}

mod.PD.complete <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEP.BWexp_1.RDS")

summary(mod.PD.complete)

#check_model(mod.PD.complete)

rm(list = ls())

```

### Figure 5 A

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEP ~ is.forest + s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
mod.PD.PC1 <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEP.BWexp_2.RDS")

summary(mod.PD.PC1)

rm(list = ls())
```

### Figure 5 B

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.RQEP ~ PC1 + s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
mod.PD.for <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEP.BWexp_3.RDS")

summary(mod.PD.for)

rm(list = ls())
```

<!-- ### Figure 5 C -->

<!-- #### Model structure -->

<!-- ```{r, echo=TRUE, eval=FALSE} -->
<!-- mgcv::gam(SES.RQEP ~ 1 + s(Longitude, Latitude, bs = "sos"), -->
<!--           family = "gaussian", -->
<!--           method = "REML", -->
<!--           data = data.wgs) -->
<!-- ``` -->

<!-- #### Summary -->

<!-- ```{r, echo = FALSE,comment=NA} -->
<!-- mod.PD.sp <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEP-exp_3.RDS") -->

<!-- summary(mod.PD.sp) -->

<!-- rm(list = ls()) -->
<!-- ``` -->

## Figure 6

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(log(SES.RQEF/SES.RQEP) ~ PC1 + is.forest + s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}

mod.rel.complete <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_rel.log.BW exp_10.RDS")

summary(mod.rel.complete)

#check_model(mod.PD.complete)

rm(list = ls())

```

### Figure 6 A

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(log(SES.RQEF/SES.RQEP) ~ is.forest + s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
mod.rel.PC1 <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_rel.log.BW exp_11.RDS")

summary(mod.rel.PC1)

rm(list = ls())
```

### Figure 6 B

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(log(SES.RQEF/SES.RQEP) ~ PC1 + s(Longitude, Latitude, bs = "sos"),
          family = "gaussian",
          method = "REML",
          data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
mod.rel.for <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_rel.log.BW exp_12.RDS")

summary(mod.rel.for)

rm(list = ls())
```


## Figure S 1

### Figure S 1 A

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(FD ~ PD + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS1A <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-RQEF-RQEP.Rds")

summary(modS1A)

#check_model(modS1A)

rm(list = ls())
```

### Figure S 1 B

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(FD ~ PD.sqrt + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS1B <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-RQEF-RQEP.sqrt.Rds")

summary(modS1B)

#check_model(modS1B)

rm(list = ls())
```

### Figure S 1 C

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.FD ~ SES.PD + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS1C <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-SES-RQEF-RQEP.Rds")

summary(modS1C)

#check_model(modS1B)

rm(list = ls())
```

### Figure S 1 D

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.FD ~ SES.PD.sqrt + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS1D <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-SES-RQEF-RQEP.sqrt.Rds")

summary(modS1D)

rm(list = ls())
```

### Figure S 1 E

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.FDis ~ SES.MPD + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS1E <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-FDis-MPD-spatial.Rds")

summary(modS1E)

rm(list = ls())
```

### Figure S 1 F

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.FD ~ SES.PD + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS1F <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-SES-RQEF-RQEP.bio.Rds")

summary(modS1F)

rm(list = ls())
```
<!-- ### Figure S 1 D -->

<!-- #### Model structure -->

<!-- ```{r, echo=TRUE, eval=FALSE} -->
<!-- mgcv::gam(FD ~ 1 + s(Longitude, Latitude, bs = "sos"), -->
<!--                     family = "gaussian", -->
<!--                     method = "REML", -->
<!--                     data = data.wgs) -->
<!-- ``` -->

<!-- #### Summary -->

<!-- ```{r, echo = FALSE,comment=NA} -->
<!-- modS1D <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-RQEF-RQEP.Smooth.Rds") -->

<!-- summary(modS1D) -->

<!-- #check_model(modS1C) -->

<!-- rm(list = ls()) -->
<!-- ``` -->

## Figure S 2

### Figure S 2 A

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(FD ~ 1 + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS2A <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-RQEF-RQEP.Smooth.Rds")

summary(modS2A)

#check_model(modS1C)

rm(list = ls())
```


### Figure S 2 B

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.FD ~ 1 + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS2B <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-SES-RQEF-smooth.Rds")

summary(modS2B)

#check_model(modS1C)

rm(list = ls())
```

### Figure S 2 C

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.FD ~ 1 + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS2C <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-SES-RQEF.bio-smooth.Rds")

summary(modS2C)

#check_model(modS1C)

rm(list = ls())
```

### Figure S 2 D

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.FD ~ 1 + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS2D <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/04.GAM-SES-RQEF.bio.weigh-smooth.Rds")

summary(modS2D)

#check_model(modS1C)

rm(list = ls())
```

## Figure S 6

### Figure S 6 A

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.FD ~ 1 + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS6A <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEF.BWexp_9.RDS")

summary(modS6A)

#check_model(modS1C)

rm(list = ls())
```


### Figure S 6 B

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(SES.PD ~ 1 + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS6B <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_SES.RQEP.BWexp_4.RDS")

summary(modS6B)

#check_model(modS1C)

rm(list = ls())
```

### Figure S 6 C

#### Model structure

```{r, echo=TRUE, eval=FALSE}
mgcv::gam(log(SES.FD/SES.PD) ~ 1 + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
modS6C <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/05.GAM_rel.log.BW exp_13.RDS")

summary(modS6C)

#check_model(modS1C)

rm(list = ls())
```


## Figure S 7

### Figure S 7 A

#### Model structure

```{r, echo=TRUE, eval=FALSE}

indices <- c(
    "PDQ",
    "FDQ.ALL",
    "FDQ.SLA",
    "FDQ.HEIGHT",
    "FDQ.ROOT",
    "FDQ.SM",
    "FDQ.LDMC",
    "FDQ.N",
    "FDQ.P",
    "FDQ.CHRO.PA")

mgcv::gam(indices[i] ~ PDQ.PA + s(Longitude, Latitude, bs = "sos"),
                    family = "gaussian",
                    method = "REML",
                    data = data.wgs)
```

#### Summary

```{r, echo = FALSE,comment=NA}
models <- readRDS("~/09.projects/2022/09.sPlot-project-FD-PD/02.data/sPlotOpen-PA-GAMS.Rds")

blanks <- rep(c(' ', '\n'),3) # This will print five blank lines between plots.

AddBreak <- function() {
for (i in blanks)
  cat(i)
}

for(i in 1:length(models)) {
  print(summary(models[[i]]))
  AddBreak()
}

rm(list = ls())
```